Fast processing of JUNGFRAU data with the Alpaka library

### 2019-07-03 Publication/manuscript.tex
Added some related works.

\documentclass[a4paper]{article}
\usepackage{authblk}
\usepackage[backend=bibtex]{biblatex}
\usepackage{hyperref}
\usepackage{txfonts}
\usepackage{titling}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[a4paper, margin={2.5cm, 1.7cm}]{geometry}
\usepackage{mwe}
\usepackage{filecontents}
\usepackage{changes}
\usepackage{lipsum}% <- For dummy text
\definechangesauthor[name={Heide}, color=blue]{Heide}
\setremarkmarkup{(#2)}

@@

\affil[1]{Helmholtz-Zentrum Dresden -- Rossendorf, Bautzner Landstra{\ss}e 400, 01328 Dresden, Germany}
\affil[2]{Technische Universit\"at Dresden, 01069 Dresden}
\affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland}

@@

The increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing.
At PSI, the Jungfrau detector (\cite{JFapplications}, \cite{JFcalibration}, \cite{JFoperation}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode. A conversion of the detector data to the energy and the number of registered photons already during data aquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space. This online data conversion algorithm has to correct each pixel value using the gain maps and the continously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}. The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data. The different modules of one detector system can be processed fully parallel, however, the data conversion itself is not data parallel. Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka~\cite{Matthes17} version was written and its performance was tested.\\

\textcolor{blue}{TODO: Related work}\\

@@

@@

>>>>

\documentclass[a4paper]{article}
\usepackage{authblk}
\usepackage[backend=bibtex]{biblatex}
\usepackage{hyperref}
\usepackage{txfonts}
\usepackage{titling}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[a4paper, margin={2.5cm, 1.7cm}]{geometry}
\usepackage{mwe}
\usepackage{filecontents}
\usepackage{changes}
\usepackage{lipsum}% <- For dummy text
%\definechangesauthor[name={Heide}, color=blue]{Heide}
%\definechangesauthor[name={Jonas}, color=pink]{Jonas}
%\setremarkmarkup{(#2)}

@@

\affil[1]{Helmholtz-Zentrum Dresden -- Rossendorf, Bautzner Landstra{\ss}e 400, 01328 Dresden, Germany}
\affil[2]{Technische Universit\"at Dresden, 01069 Dresden, Germany}
\affil[3]{Paul Scherrer Institut, 5232 Villigen-PSI,  Switzerland}

@@

The increasing data rates during FEL experiments require dedicated detectors as well as advanced methods for fast data processing.
At PSI, the Jungfrau detector (\cite{JFapplications}, \cite{JFcalibration}, \cite{JFoperation}, \cite{Mozzanica_2016}) was developed which is capable due to a gain switching scheme to register single photons as well as photon bunches in a charge integration mode. A conversion of the detector data to the energy and the number of registered photons already during data acquisition is beneficial for the analysis of the measurement and furthermore for the reduction of storage space. This online data conversion algorithm has to correct each pixel value using the gain maps and the continuously updated pedestal maps following a lab-based calibration  procedure~\cite{JFcalibration}. The summation of a couple of frames and the cluster identification in the data maps further reduce the amount of data. The different modules of one detector system can be processed fully parallel, however, the data conversion itself is not data parallel. Therefore, and due to the need for an advanced parallel implementation which is independent of the computing hardware, an Alpaka \cite{Matthes17} version was written and its performance was tested.\\

\textcolor{blue}{TODO: Related work}\\
\textcolor{pink}{
common preprocessing steps:
 * offset correction \& flat field correction [DSSC \cite{moch2014calibration}, JF \cite{JFcalibration}, EIGER\cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]]
 * gain correction [DSSC \cite{moch2014calibration}, EIGER \cite{tinti2018electron} / \cite{dinapoli2013eiger}, MOENCH \cite{cartier2016micrometer}]
 * dynamic gain switching [AGIPD \cite{allahgholi2015agipd}, GOTTHARD \cite{mozzanica2012gotthard}, JF \cite{JFcalibration}, [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
 * pixel sorting (on FPGA)[AGPID \cite{becker2013high}]
 * discard uninteresting events (in readeout circuit)[DSSC \cite{erdinger2012dssc}]
 * photon counting [MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
 * charge summing / interpolation (to counter charge sharing; on ASIC)[MEDIPIX3RX \cite{ballabriga2013medipix3rx}]
 * cluster finder [MOENCH \cite{cartier2016micrometer}]
 * sub pixel interpolation [MOENCH \cite{cartier2016micrometer}]}\\
%TODO: common calibration and preprocessing steps / online preprocessing -> FPGAs
%TODO: offline preprocessing -> XDS, ...
%TODO: further analysis -> crystallography
%TODO: related acceleration libraries

@@

requirements:
 - realtime data processing (at least 1 detector at 2.5 kHz)
 - simple and extensible design
 - portable to other architectures (CPU OpenMP, TBB, ...) with minimal code changes

optional requirements:
 - easily adaptable to similar detectors (e.g. MÖNCH)

design overview:
 - wrap for most important alpaka functions (for ease of use)
 - templated for maximal performance and configurability
 - stream processing of data: upload data to GPU -> process -> download again
 - multiple detectors and multiple GPUs usable in parallel

@@

benchmarked with google benchmark
